---
title: "writing_functions"
author: "rl3411"
date: "2023-10-26"
output: github_document
---

```{r}
library(tidyverse)
library(rvest)
```

Set seed for reproducibility

```{r}
set.seed(12345)
```

### Z score function

 Z scores subtract the mean and divide by the sd.
 
```{r}
x_vec = rnorm(20, mean = 5, sd = 0.3)
```

Compute Z scores for `x_vec`.

```{r}
(x_vec - mean(x_vec)) / sd(x_vec)
```

 Write a function to do this!
 
```{r}
z_score = function(x) {
  z = (x - mean(x)) / sd(x)
  z
}
```

Check that this works 
```{r}
z_score(x = x_vec)
```
 
 Keep checking.
 
```{r, eval = F}
z_score(x=3)
z_score(c("my", "name", "is", "jeff"))
z_score(c(TRUE,TRUE,FALSE,FALSE))
z_score(iris)
```

 
 Now improve the function
 
```{r}
z_score = function(x) {
  if(!is.numeric(x)) {
    stop("Argument should be numbers")
  } else if (length(x) < 2) {
    stop("You need at least 2 numbers to get z scores")
  }
  z = (x - mean(x)) / sd(x)
  z
}
```

### Multiple outputs

Write a function that returns the mean and sd from a sample of numbers

```{r}
mean_and_sd = function(x) {
  if(!is.numeric(x)) {
    stop("Argument should be numbers")
  } else if (length(x) < 2) {
    stop("You need at least 2 numbers to get z scores")
  }
  mean_x = mean(x) 
  sd_x = sd(x)
  tibble(
    mean = mean_x,
    sd = sd_x
  )
}
```

Double check

```{r}
mean_and_sd(x_vec)
```

```{r}
mean_and_sd = function(x) {
  mean_x = mean(x) 
  sd_x = sd(x)
  tibble(
    mean = mean_x,
    sd = sd_x
  )
}
```

### Start getting means and sds

```{r}
x_vec = rnorm(n = 30, mean = 5, sd = 0.5)

tibble(
  mean = mean(x_vec),
  sd = sd(x_vec)
)
```

Let's write a function that uses `n`, a true mean and true SD as inputs

```{r}
sim_mean_sd = function(n_obs, mu, sigma) {
  x_vec = rnorm(n=n_obs, mean = 5, sd = 0.5) 
  tibble(
    mean = mean(x_vec), 
    sd = sd(x_vec)
  )
}

sim_mean_sd(n_obs = 30, mu = 5, sigma = 0.5)
sim_mean_sd(12, 24, 4)
```





