---
title: "writing_functions"
author: "rl3411"
date: "2023-10-26"
output: github_document
---

```{r}
library(tidyverse)
library(rvest)
```

Set seed for reproducibility

```{r}
set.seed(12345)
```

### Z score function

 Z scores subtract the mean and divide by the sd.
 
```{r}
x_vec = rnorm(20, mean = 5, sd = 0.3)
```

Compute Z scores for `x_vec`.

```{r}
(x_vec - mean(x_vec)) / sd(x_vec)
```

 Write a function to do this!
 
```{r}
z_score = function(x) {
  z = (x - mean(x)) / sd(x)
  z
}
```

Check that this works 
```{r}
z_score(x = x_vec)
```
 
 Keep checking.
 
```{r, eval = F}
z_score(x=3)
z_score(c("my", "name", "is", "jeff"))
z_score(c(TRUE,TRUE,FALSE,FALSE))
z_score(iris)
```

 
 Now improve the function
 
```{r}
z_score = function(x) {
  if(!is.numeric(x)) {
    stop("Argument should be numbers")
  } else if (length(x) < 2) {
    stop("You need at least 2 numbers to get z scores")
  }
  z = (x - mean(x)) / sd(x)
  z
}
```

### Multiple outputs

Write a function that returns the mean and sd from a sample of numbers

```{r}
mean_and_sd = function(x) {
  if(!is.numeric(x)) {
    stop("Argument should be numbers")
  } else if (length(x) < 2) {
    stop("You need at least 2 numbers to get z scores")
  }
  mean_x = mean(x) 
  sd_x = sd(x)
  tibble(
    mean = mean_x,
    sd = sd_x
  )
}
```

Double check

```{r}
mean_and_sd(x_vec)
```

```{r}
mean_and_sd = function(x) {
  mean_x = mean(x) 
  sd_x = sd(x)
  tibble(
    mean = mean_x,
    sd = sd_x
  )
}
```

### Start getting means and sds

```{r}
x_vec = rnorm(n = 30, mean = 5, sd = 0.5)

tibble(
  mean = mean(x_vec),
  sd = sd(x_vec)
)
```

Let's write a function that uses `n`, a true mean and true SD as inputs

```{r}
sim_mean_sd = function(n_obs, mu, sigma) {
  x_vec = rnorm(n=n_obs, mean = 5, sd = 0.5) 
  tibble(
    mean = mean(x_vec), 
    sd = sd(x_vec)
  )
}

sim_mean_sd(n_obs = 30, mu = 5, sigma = 0.5)
sim_mean_sd(12, 24, 4)
```

### Load LoTR data

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") |>
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") |>
  mutate(movie = "return_king")

lotr_tidy = bind_rows(fellowship_ring, two_towers, return_king) |>
  janitor::clean_names() |>
  gather(key = sex, value = words, female:male) |>
  mutate(race = str_to_lower(race)) |> 
  select(movie, everything()) 
```

``` {r}
lotr_load_and_tidy = function(path = "data/LotR_Words.xlsx", range, movie_name) {
  df = readxl::read_excel(path, range = range) |>
    janitor::clean_names() |>
    pivot_longer(
      female:male,
      names_to = "sex",
      values_to = "words") |>
    mutate(race = str_to_lower(race),
           movie = movie_name) |> 
    select(movie, everything())
  df
}

lotr_tidy = 
  bind_rows(
    lotr_load_and_tidy("data/LotR_Words.xlsx", "B3:D6", "fellowship_ring"),
    lotr_load_and_tidy("data/LotR_Words.xlsx", "F3:H6", "two_towers"),
    lotr_load_and_tidy("data/LotR_Words.xlsx", "J3:L6", "return_king")) |>
  select(movie, everything())
```

### NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

Instead of changing the table name, html and selecting nth line, can write
a function instead.

```{r}
nsduh_table <- function(html, table_num, table_name) {
  
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent),
      name = table_name) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
}
```

Check function

```{r}
nsduh_results = 
  bind_rows(
    nsduh_table(nsduh_html, 1, "marj_one_year"),
    nsduh_table(nsduh_html, 4, "cocaine_one_year"),
    nsduh_table(nsduh_html, 5, "heroin_one_year")
  )
```










